---
title: "Ocean R"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Install system libraries

```{bash}
sudo apt-get install libnetcdf-dev
```

# Install required packages

```{r}
install.packages(c("getPass","ini","aws.s3"))
```

# Notebook functions, libraries and environmental variables

```{r}
# %% Libraries
library(getPass)

# %% Functions
get_app_config_dir <- function(appname) {
  app_rel_path <- file.path(".config", appname, paste0(appname, ".conf"))
  app_system_path <- file.path(path.expand("~"), app_rel_path)
  return(app_system_path)
}

# %% Env Var
appname            <- "oceananalysis"
server_config_name <- "default"
```

# Setting Abyssio configuration

```{r}
# %% Main
app_system_path <- get_app_config_dir(appname = appname)
dir.create(dirname(app_system_path), recursive = TRUE, showWarnings = FALSE)

write_file <- FALSE

if (file.exists(app_system_path)) {
  answer_options <- c("y", "n")
  repeat {
    answer <- readline(prompt = paste0(
      app_system_path,
      " already exists, do you want to overwrite? [",
      paste(answer_options, collapse = ", "), "]: "
    ))
    if (answer == answer_options[1]) {
      write_file <- TRUE
      break
    } else if (answer == answer_options[2]) {
      write_file <- FALSE
      break
    } else {
      cat("Answer was not one of the available options, please use either of the answer options:",
          paste(answer_options, collapse = ", "), "\n")
    }
  }
} else {
  write_file <- TRUE
}

if (write_file) {
  profile    <- "default"
  access_key <- readline(prompt = "Enter access key: ")
  secret_key <- getPass("Enter secret key: ")

  config_data <- paste0(
    "[", profile, "]\n",
    "access_key = ", access_key, "\n",
    "secret_key = ", secret_key, "\n",
    "endpoint = 145.38.204.177:9000\n",
    "secure = False\n"
  )

  cat("Writing configuration file to", app_system_path, "\n")
  writeLines(config_data, con = app_system_path)
  cat("Done!\n")
}
```

# Testing connection

```{r}
library(aws.s3)
library(ini)

# %% Main
app_system_path <- get_app_config_dir(appname = appname)
config          <- read.ini(app_system_path)

# Set credentials as environment variables (used by aws.s3 internally)
Sys.setenv(
  "AWS_ACCESS_KEY_ID"     = trimws(config[[server_config_name]][["access_key"]]),
  "AWS_SECRET_ACCESS_KEY" = trimws(config[[server_config_name]][["secret_key"]]),
  "AWS_S3_ENDPOINT" = trimws(config[[server_config_name]][["endpoint"]]))
options("cloudyr.aws.default_region" = "")
buckets <- bucketlist(use_https=F)
cat(buckets$Bucket, sep = "\n")
```

# Examples using Abyssio

## Setting your connection

```{r}
# Read configuration
app_system_path <- get_app_config_dir(appname = appname)
config          <- read.ini(app_system_path)

# Set credentials as environment variables (used by aws.s3 internally)
Sys.setenv(
  "AWS_ACCESS_KEY_ID"     = trimws(config[[server_config_name]][["access_key"]]),
  "AWS_SECRET_ACCESS_KEY" = trimws(config[[server_config_name]][["secret_key"]]),
  "AWS_S3_ENDPOINT" = trimws(config[[server_config_name]][["endpoint"]]))
options("cloudyr.aws.default_region" = "")
```

## Downloading data

### Downloading a single file

```{r}
# Downloading a single file

# %% Env var
# Server location information
bucket_name <- "input"
prefix      <- "netcdf"
file_name   <- "CO2f_fesom_historical_19900101.nc"
object_name <- paste(prefix, file_name, sep = "/")

# Local location information
local_download_directory <- "downloads"
local_download_path      <- file.path(path.expand("~"), local_download_directory)
local_file_path          <- file.path(local_download_path, file_name)

dir.create(local_download_path, recursive = TRUE, showWarnings = FALSE)

# %% Main
# Downloading the file
save_object(
  object   = object_name,
  bucket   = bucket_name,
  file     = local_file_path,
  use_https = F
)
```

### Downloading a full prefix (directory)

```{r}
# Downloading a full prefix non-recursively

# %% Env var
# Server location information
bucket_name <- "input"
prefix      <- "netcdf/"  # Trailing slash to look into prefix

# Local location information
local_download_directory <- "downloads"
local_download_path      <- file.path(path.expand("~"), local_download_directory)

dir.create(local_download_path, recursive = TRUE, showWarnings = FALSE)

# %% Main
stime <- as.numeric(Sys.time())  # Seconds since Unix epoch

# List objects in prefix (non-recursive)
objects <- get_bucket(
  bucket    = bucket_name,
  prefix    = prefix,
  delimiter = "/",  # delimiter prevents recursion, equivalent to recursive = False,
  use_https = F
)

# Track stats
nobjs            <- 0
bytes_downloaded <- 0

# Download the objects
for (obj in objects) {
  object_name     <- obj[["Key"]]
  local_file_path <- file.path(local_download_path, object_name)

  # Ensure subdirectory exists
  dir.create(dirname(local_file_path), recursive = TRUE, showWarnings = FALSE)

  save_object(
    object   = object_name,
    bucket   = bucket_name,
    file     = local_file_path,
    use_https = F
  )

  nobjs            <- nobjs + 1
  bytes_downloaded <- bytes_downloaded + as.numeric(obj[["Size"]])
}

# Parse and print stats
mbytes_downloaded <- bytes_downloaded / (1024^2)
elapsed           <- as.numeric(Sys.time()) - stime
mbytes_avg_xfer   <- mbytes_downloaded / elapsed

cat(sprintf(
  "Downloaded %d files, totalling %.2f MB in %.2fs with an average speed of %.2f MB/s\n",
  nobjs, mbytes_downloaded, elapsed, mbytes_avg_xfer
))
```

## Uploading data

### Uploading a single file

```{r}
# Upload a single test file
testfile_path <- file.path(path.expand("~"), "testfile.txt")
writeLines("Hello World.", testfile_path)  # Uncomment to create the test file

bucket_name <- "personal"
prefix      <- "analysis/output"
access_key  <- trimws(config[[server_config_name]][["access_key"]])
object_name <- paste(access_key, prefix, basename(testfile_path), sep = "/")

put_object(
  file     = testfile_path,
  object   = object_name,
  bucket   = bucket_name,
  use_https = F
)
```

### Uploading a directory

```{r}
# Create a test directory with files and upload them all
test_directory <- file.path(path.expand("~"), "test_dir")
dir.create(test_directory, recursive = TRUE, showWarnings = FALSE)

for (i in 0:9) {
  writeLines(paste0("This is file ", i),
             file.path(test_directory, paste0("file_", i, ".txt")))
}

# Upload the data
bucket_name <- "personal"
prefix      <- paste0("analysis/", basename(test_directory))
access_key  <- trimws(config[[server_config_name]][["access_key"]])
file_list   <- list.files(test_directory, pattern = "\\.txt$", full.names = TRUE)

for (file_path in file_list) {
  object_name <- paste(access_key, prefix, basename(file_path), sep = "/")
  put_object(
    file     = file_path,
    object   = object_name,
    bucket   = bucket_name,
    use_https = F
  )
}
```
